Portfolios of Subgraph Isomorphism Algorithms

### 2015-12-17 paper/paper.tex
Model analysis and more conclusion

@@

We demonstrated that our algorithm selection approach achieves substantial
performance improvements over the single algorithm that has the best performance
on our benchmark set. We showed that combining an algorithm selection approach
with a new incomplete variant of \LAD that is able to detect inconsistencies
and a presolver boosts performance even further.

>>>>

\subsection{Analysis of Features Used by the Model}

Analysing the final model, we saw that the most important features were, in order:

\begin{itemize}
    \item The proportion of target vertices which are at least distance 3
        apart. We introduced this feature expecting it to be helpful in
        distinguishing between \Glasgow variants: if few vertices are far
        apart, longer paths are unlikely to be useful. However, in practice
        this feature also gives a rough indication of how sparse the graph
        is---locally all-different filtering is weak and expensive on dense
        graphs, and the feature turned out to be helpful for selecting between
        \Glasgow and \LAD variants too;

    \item The mean degree of the target graph;
    \item The maximum degree of the pattern graph;
    \item The number of values removed during \IncompleteLAD presolving;
    \item And the mean degree of the pattern graph, the number of vertices in
        the pattern graph, and the number of vertices in the target graph,
        which were used once each.
\end{itemize}

\noindent This confirms that our decision to consider both the pattern graph
and the target graph for feature extraction was correct, and shows that basic
graph properties do affect the performance of sophisticated filtering
algorithms in reasonably simple ways.

\subsection{Analysis of \PathLAD versus \GlasgowTwo}

In an attempt to gain insight into the behaviour of the algorithms, we also
investigated a simpler setting comparing just \PathLAD and \GlasgowTwo. This
pair is of particular interest because they are the best ``medium-case''
algorithms that use strong and weak filtering during search respectively. We
used machine learning techniques to train a simple, human-understandable model
which is able to distinguish these cases with reasonable accuracy. The model
produced uses four rules:

\begin{enumerate}
    \item If \IncompleteLAD removes at least \SI{28}{\percent} of the pairs,
        and at least \SI{94}{\percent} of the values from at least one domain,
        then pick \PathLAD.
    \item If the target has at least $610$ vertices, and if the maximum
        distance between any two pattern vertices is at most $8$, and if the
        pattern is not regular, and if the time taken to compute the
        distance-based features on the target graph is no more than
        \SI{1277}{\ms}, then pick \PathLAD.
    \item If \IncompleteLAD filtering removed at least \SI{6}{\percent} of
        possible pairs, and if less than \SI{85}{\percent} of the pattern
        vertices are within distance 2 of each other, then pick \PathLAD.
    \item Otherwise, pick \GlasgowTwo.
\end{enumerate}

\noindent
The first and third rules here intuitively make sense: if \IncompleteLAD
filtering does well, it is likely that continuing with this kind of filtering
during search will be successful. The third rule also excludes using \PathLAD
on very dense pattern graphs, where locally all-different filtering is
expensive and weak. The second rule is less obvious: \PathLAD filtering is weak
on regular graphs, so this part of the rule is unsurprising, but the remainder
of the rule appears to be a round-about way of excluding large, dense target
graphs. Still, this model suggests that it could be worth exploring
\emph{dynamically} enabling or disabling locally all-different filtering during
search, based upon very simple features which could be recomputed as search
progresses and conditions change.

@@

We demonstrated that our algorithm selection approach achieves substantial
performance improvements over the single algorithm that has the best performance
on our benchmark set. We showed that combining an algorithm selection approach
with a new incomplete variant of \LAD that is able to detect inconsistencies
and a presolver boosts performance even further. Finally, we discussed how
insights from machine learning could feed back into algorithm development.

We have yet to investigate parallel portfolios: as well as the possibility of
running multiple single-threaded solvers simultaneously, the \Glasgow
algorithms have a multi-core parallel implementation available, which can use a
configurable number of threads. Nor have we investigated variants of the
subgraph isomorphism problem---it would be interesting to see, for example,
whether similar features are important when dealing with induced or directed
isomorphisms.
